Circumnuclear Media of Quiescent Supermassive Black Holes

### 2015-06-19 ms.tex
referee 2

\begin{document}
\title{Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes}
\maketitle

@@

Understanding why most SMBHs appear to be inactive requires
characterizing their gaseous environments.  Gas near the SMBH sphere
of influence, hereafter denoted the `circumnuclear medium' (CNM),
controls the mass accretion rate, $\dot{M}$.  The accretion rate in
turn determines the SMBH luminosity and the feedback of its energy and
momentum output on larger scales.  Dense gas in the nucleus may lead to runaway cooling, resulting in bursty episodes of
star formation and AGN activity (e.g.~\citealt{Ciotti&Ostriker07}).

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other
properties of the nucleus informs key questions related to the
co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{HeckmanBest:2014a}).  In the
low redshift universe, SMBH growth is dominated by low mass black
holes, $M_{\bullet} \lesssim 10^{8}M_{\odot}$
(e.g.~\citealt{Heckman+04}), a fact often attributed to the trend of
`cosmic down-sizing' resulting from hierarchical structure growth
(e.g, \citealt{Gallo+08}).  However, the physical processes by which typical
low mass black holes accrete could in principle be distinct from those
operating at higher SMBH masses, or those in AGN.  Of key importance is whether SMBHs grow primarily by the
accretion of gas fed in directly from galactic or extra-galactic
scales, or whether significant growth can result also from local
stellar mass loss in the nuclear region.

@@

\citet{Ho:2009a} determines the SMBH accretion rates in a sample of
early-type galaxies by using X-ray observations to determine the Bondi
accretion rate and based on estimates for the mass loss rate of
evolved stars.  Both methods lead him to conclude that the available
gas reservoir is more than sufficient to power the observed
low-luminosity AGN, assuming the standard $\sim$ 10 per cent radiative
efficiency for thin disk accretion.  Several lines of evidence now
suggest that low-luminosity AGN result from accretion proceeding in a
radiatively inefficient mode (\citealt{Yuan&Narayan14}), due either to
the advection of gravitationally-released energy across the SMBH
horizon (e.g.~\citealt{Narayan&Yi95}) or due to disk outflows, which
reduce the efficiency with which the inflowing gas ultimately reaches
the SMBH (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

Another approach to determine the SMBH accretion rate, which we adopt
in this paper, is to directly calculate the density, velocity and
temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected into the nuclear environment via
stellar winds, while energy is input from several sources including
stellar winds, supernovae, and AGN feedback
(\citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).
Unlike previous works, which focused primarily on modeling individual
galaxies, here we model the CNM properties across a
representative range of galaxy properties, including different SMBH
masses, stellar density profiles, and star formation histories (SFHs).

Previous studies, employing multi-dimensional numerical hydrodynamics and
including variety of (parameterized) physical effects, have focused on
massive elliptical galaxies (e.g.~\citealt{Ciotti&Ostriker07};
\citealt{Ciotti+10}).  These works show the periodic
development of cooling instabilities on galactic scales, which
temporarily increase the gas accretion rate towards the nucleus until
feedback becomes strong enough to shut off the flow and halt
SMBH growth.

@@

In the presence of strong heating, one-dimensional steady-state flow
is characterized by an inflow-outflow structure, with a critical
radius known as the``stagnation radius" $\rs$ where the radial
velocity passes through zero.  Mass loss from stars interior to the
stagnation radius is accreted, while that outside $\rs$ is unbound in
an outflow from the nucleus.  The stagnation radius, rather than the
Bondi radius, thus controls the SMBH accretion rate (although we will
show that $\rs$ usually resides near the nominal Bondi radius).  When
heating is sufficiently weak, however, the stagnation radius may move
to much larger radii or not exist at all, in which case nearly the
entire ISM of the galaxy is flowing towards the SMBH, i.e. a ``cooling
flow".  However, the hydrostatic nature of gas near the stagnation
radius also renders the CNM at this location particularly susceptible
to {\it local} thermal instabilities, the outcome of which could well
be distinct from the development of a cooling flow.

@@

\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\label{sec:gal_model}
\citet{LauerFaber+:2007a} use Hubble Space Telescope WFPC2 imaging to
measure the radial surface brightness profiles for hundreds of nearby
early type galaxies. The measured profile is fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b},
\end{equation}
i.e., a broken power law that transitions from an inner power law
slope, $\Gamma$, to an outer power law slope, $\beta$, at a break
radius, $\rb$.  If the stellar population is spherically symmetric,
then this corresponds to a 3D stellar density $\rhostar \propto
r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\propto r^{-1-\beta}$ for
$r \gg \rb$.  We can write the deprojected stellar density
approximately (formally, this is the $\alpha \rightarrow \infty$
limit) as
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star|_{\rinf} \left(r/\rinf\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star|_{r_b} \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star|_{\rinf}$ is the stellar density at the radius of
the black hole sphere of influence\footnote{This approximate
  expression agrees surprisingly well with the mean empirical scaling
  relation for $r_{\rm inf}(M_\bullet)$ calculated in Appendix C of
  \citet{Stone&Metzger15}, although we note that this relation has
  significant scatter. Also, we note that the  scaling is somewhat
  different for cores and cusps. Fixing the power law slope,
  $\rinf\simeq 25 (8) \Mbheight^{0.6}$ for core (cusp)
  galaxies. Unless otherwise noted we will use the single relation in
  equation~\eqref{eq:rsoi}},
\begin{equation}
  \rinf \simeq G \Mbh/\sigma_{\bullet}^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\end{equation}
where $M_{\bullet,8} \equiv M_{\bullet}/10^{8}M_{\odot}$ and the
second equality in (\ref{eq:rsoi}) employs the $\Mbh-\sigma$
relationship of \citet{McConnellMa+:2011a},
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma_{\bullet}}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
Though of questionable validity for low mass black holes (e.g.,
\citealt{Greene+2010, Kormendy+2013}), we adopt equation
(\ref{eq:Msigma}) for lack of an obviously better alternative.

@@

@@

The large scale inflow rate towards the SMBH is given by the total
mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]),
\begin{eqnarray}
  \dot{M} &=& 4\pi \int_{0}^{r_{s}}q r^{2}dr = \frac{\eta
    M_{\star}|_{r_{\rm s}}}{\th} = \frac{\eta \Mbh (\rs/\rinf)^{2-\Gamma}}{\th} \nonumber \\
  &\approx&
  \begin{cases}
    4.5 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    3.2 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}            
\end{eqnarray}
where we have assumed $v_{w} \gg \sigma_0$ by adopting equation
(\ref{eq:stag_simple}) for $r_s$.  The resulting Eddington ratio is
given by
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    2.0 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core}, \\
    1.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
where $\dot{M}_{\rm edd} = 2.2M_{\bullet,8}M_{\odot}$ yr$^{-1}$ is the
Eddington accretion rate, assuming a radiative efficiency of ten per
cent.  Note the sensitive dependence of the inflow rate on the wind
heating rate.  As will be discussed later,
equation~\eqref{eq:eddr_analytic} is the radial mass inflow rate on
relatively large scales and does not include outflows from the SMBH
accretion disk (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}),
which may significantly reduce the fraction of $\dot{M}$ that actually
reaches the SMBH. Thus we distinguish between the large scale inflow
rate, $\dot{M}$, and the accretion rate onto the black hole,
$\dot{M}_{\bullet}$.

@@

\begin{align}
  \frac{\dot{M}_{\rm TI}}{\dot{M}_{\rm edd}} \simeq \begin{cases}
    2\times 10^{-4}\eta_{0.02}^{0.47}M_{\bullet,8}^{0.36} &, \text{core}, \\
    8\times 10^{-5} \eta_{0.02}^{0.59} M_{\bullet,8}^{0.28}    &, \text{cusp},\\
  \end{cases}
  \label{eq:Mdotmax}
\end{align}
where the cusp case is derived using equation (\ref{eq:eddr_analytic})
to calculate $\dot{M}(v_{\rm TI})$, and the core case corresponds to
the more stringent $\zeta > \zeta_c$ criterion, calculated again using
the mean $r_{b}/r_{\rm inf}$ relation from \citet{LauerFaber+:2007a}.

What we describe above as ``thermal instability" may in practice
simply indicate an abruption transition from a steady inflow-outflow
solution to a global cooling flow, as opposed a true thermal
instability.  In the former case the stagnation radius diverges to
large radii, increasing the density in the inner parts of the flow,
which increases cooling and creates a large inflow of cold gas towards
the nucleus.  A true thermal instability would likely result in a
portion of the hot ISM condensing into cold clouds, a situation which
may or may not be present in a cooling flow.  In this paper we do not
distinguish between these possibilities, although both are likely
present at some level.  Finally, note that if the CNM were to
``regulate" itself to a state of local marginal thermal instability
(as has previously been invoked on cluster scales; e.g.,
\citealt{Voit+15}), then equation (\ref{eq:Mdotmax}) might naturally
reflect the characteristic mass fall-out rate and concomitant star
formation rate.

@@

Key properties of the flow, such as the mass inflow rate and the
likelihood of thermal instability depend sensitively on the assumed
heating rate $\propto qv_{w}^{2}$.  This section summarizes individual
heating sources, providing estimates of the value of the effective
wind velocity
\begin{align}
  v_{w} = \sqrt{\frac{2 t_h \dot{e}}{\eta \rhostar}}
  \label{eq:vw_eff}
\end{align}
for each, where $\dot{e}$ is volumetric heating rate.

@@

where we implicitly the different sources of energy injection mix
efficiently. In reality, this may not be the case as slower velocity
sources cool and form high density structure due to external pressure
from faster winds \citep{Cuadra+2005}.

@@

Perhaps a more readily addressable test is whether a steady-state
accretion picture developed in this work can account for the the large
scatter, typically of $2-3$ orders of magnitude, in $L_{X}/L_{\rm
  edd}$ at fixed $M_{\bullet}$.  Scatter could result from the strong
sensitivity of the accretion rate to the stellar wind velocity and
mass loss parameter, $\dot{M}/\dot{M}_{\rm edd} \propto
v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively
(Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).  When
combined with the significant dependence of $v_w$ on stochastic
intermittency in the star formation history (Appendix
\ref{fig:NickPlot2}), this can lead to order of magnitude differences
in $\dot{M}$.  However, we note that $v_w$ is not expected to vary by
more than a factor of a few for a thermally-stable flow, limiting the
allowed variation of $\dot{M}/\dot{M}_{\rm edd}$.  Variations in
$L_{X}/L_{\rm edd}$ could also result from differences in angular
momentum of the infalling gas from galaxy to galaxy, resulting in
differences in the fraction of the gas lost to outflows.  Also
potentially contributing is the order of magnitude difference, at
fixed $v_w$ and $M_{\bullet}$, between $\dot{M}$ for core and cusp
galaxies (Fig.~\ref{fig:bh_growth}).  Differences in $\dot{M}$ are
augmented by the theoretical expectation that $L_{\rm X} \propto
\dot{M}^{2} $ for radiatively inefficient flows.

@@

  \item The location of the stagnation radius, and hence the SMBH
    accretion rate, is a sensitive function of the gas heating rate
    $\propto v_w^{2}$.  Quantitatively, $\rs\propto v_w^{-2}$,
    implying that $\Mdot\propto v_w^{-3.8 (-2.4)}$ for fiducial core
    (cusp) galaxies.  However, $\Mdot$ can increase even more rapidly
    with decreasing $v_w$ for two reasons: (1) when $v_w$ becomes
    comparable to the stellar velocity dispersion ($\zeta < \zeta_c$;
    eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar
    bulge and cannot produce a flow with a positive kinetic energy
    interior to the stellar break radius, as we show in Appendix
    \ref{app:be}; (2) For low $v_w$, gas near the stagnation radius
    becomes thermally unstable, which based on the results of previous
    numerical studies (e.g.,\citealt{Ciotti+10}) is instead likely to
    result in a burst of high accretion ($\S\ref{sec:cycle}$).

@@

>>>>

\begin{document}
\title{Circumnuclear Media of Quiescent Supermassive Black Holes}
\maketitle

@@

Understanding why most SMBHs appear to be inactive requires
characterizing their gaseous environments.  Gas near the SMBH sphere
of influence, hereafter denoted the `circumnuclear medium' (CNM),
controls the mass accretion rate, $\dot{M}_{\bullet}$.  The accretion
rate in turn determines the SMBH luminosity and the feedback of its
energy and momentum output on larger scales.  Dense gas in the nucleus
may lead to runaway cooling, resulting in bursty episodes of star
formation and AGN activity (e.g.~\citealt{Ciotti&Ostriker07}).

Knowledge of how $\dot{M}_{\bullet}$ depends on the SMBH mass, $\Mbh$,
and other properties of the nucleus informs key questions related to
the co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{HeckmanBest:2014a}).  In the
low redshift universe, SMBH growth is dominated by low mass black
holes, $M_{\bullet} \lesssim 10^{8}M_{\odot}$
(e.g.~\citealt{Heckman+04}), a fact often attributed to the trend of
`cosmic down-sizing' resulting from hierarchical structure growth
(e.g, \citealt{Gallo+08}).  However, the physical processes by which
typical low mass black holes accrete could in principle be distinct
from those operating at higher SMBH masses, or those in AGN.  Of key
importance is whether SMBHs grow primarily by the accretion of gas fed
in directly from galactic or extra-galactic scales, or whether
significant growth can result also from local stellar mass loss in the
nuclear region.

@@

{\bf AG need to introduce concept of inflow rate here}
\citet{Ho:2009a} determines the inflow rates in a sample of early-type
galaxies by using X-ray observations to determine the Bondi accretion
rate and based on estimates for the mass loss rate of evolved stars.
Both methods lead him to conclude that the available gas reservoir is
more than sufficient to power the observed low-luminosity AGN,
assuming the standard $\sim$ 10 per cent radiative efficiency for thin
disk accretion.  Several lines of evidence now suggest that
low-luminosity AGN result from accretion proceeding in a radiatively
inefficient mode (\citealt{Yuan&Narayan14}), due either to the
advection of gravitationally-released energy across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to disk outflows, which reduce
the efficiency with which the inflowing gas ultimately reaches the
SMBH (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

Another approach to determine the inflow rates, which we adopt
in this paper, is to directly calculate the density, velocity and
temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected into the nuclear environment via
stellar winds, while energy is input from several sources including
stellar winds, supernovae, and AGN feedback
(\citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).
Unlike previous works, which focused primarily on modeling individual
galaxies, here we model the CNM properties across a
representative range of galaxy properties, including different SMBH
masses, stellar density profiles, and star formation histories (SFHs).

Previous studies, employing multi-dimensional numerical hydrodynamics and
including variety of (parameterized) physical effects, have focused on
massive elliptical galaxies (e.g.~\citealt{Ciotti&Ostriker07};
\citealt{Ciotti+10}).  These works show the periodic
development of cooling instabilities on galactic scales, which
temporarily increase the gas inflow rate towards the nucleus until
feedback becomes strong enough to shut off the flow and halt
SMBH growth.

@@

In the presence of strong heating, one-dimensional steady-state flow
is characterized by an inflow-outflow structure, with a critical
radius known as the``stagnation radius" $\rs$ where the radial
velocity passes through zero.  Mass loss from stars interior to the
stagnation radius is accreted, while that outside $\rs$ is unbound in
an outflow from the nucleus.  The stagnation radius, rather than the
Bondi radius, thus controls the inflow rate (although we will
show that $\rs$ usually resides near the nominal Bondi radius).  When
heating is sufficiently weak, however, the stagnation radius may move
to much larger radii or not exist at all, in which case nearly the
entire ISM of the galaxy is flowing towards the SMBH, i.e. a ``cooling
flow".  However, the hydrostatic nature of gas near the stagnation
radius also renders the CNM at this location particularly susceptible
to {\it local} thermal instabilities, the outcome of which could well
be distinct from the development of a cooling flow.

@@

\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\label{sec:gal_model}
\citet{LauerFaber+:2007a} use Hubble Space Telescope WFPC2 imaging to
measure the radial surface brightness profiles for hundreds of nearby
early type galaxies. The measured profile is fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b},
\end{equation}
i.e., a broken power law that transitions from an inner power law
slope, $\Gamma$, to an outer power law slope, $\beta$, at a break
radius, $\rb$.  If the stellar population is spherically symmetric,
then this corresponds to a 3D stellar density $\rhostar \propto
r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\propto r^{-1-\beta}$ for
$r \gg \rb$.  We can write the deprojected stellar density
approximately (formally, this is the $\alpha \rightarrow \infty$
limit) as
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star|_{\rinf} \left(r/\rinf\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star|_{r_b} \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star|_{\rinf}$ is the stellar density at the radius of
the black hole sphere of influence\footnote{This approximate
  expression agrees surprisingly well with the mean empirical scaling
  relation for $r_{\rm inf}(M_\bullet)$ calculated in Appendix C of
  \citet{Stone&Metzger15}, although we note that this relation has
  significant scatter. Also, we note that the  scaling is somewhat
  different for cores and cusps. Fixing the power law slope,
  $\rinf\simeq 25 (8) \Mbheight^{0.6}$ for core (cusp)
  galaxies. Unless otherwise noted we will use the single relation in
  equation~\eqref{eq:rsoi}},
\begin{equation}
  \rinf \simeq G \Mbh/\sigma_{\bullet}^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\end{equation}
where $M_{\bullet,8} \equiv M_{\bullet}/10^{8}M_{\odot}$ and the
second equality in (\ref{eq:rsoi}) employs the $\Mbh-\sigma$
relationship of \citet{McConnellMa+:2011a},
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma_{\bullet}}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
This may be of questionable validity for low mass black holes (e.g.,
\citealt{Greene+2010, Kormendy+2013}). Also, several of the black hole
masses used in \citealt{McConnellMa+:2011a} were underestimated
\citep{Kormendy+2013}.

@@

Our model does not take into account complications such as more
complicated geometries or the discrete nature of real stars
\citep{Cuadra+2006, Cuadra+2008}. In the case of Sgr A* these effects
reduce the time average inflow rate to $\sim10^{-6} \Msun \pyear$ --
an order of magnitude less than a 1D spherical model
\citep{Cuadra+2006}.  Motions of individual stars can produce ~order
of magnitude spikes in the accretion rate \citep{Cuadra+2008}.

@@

The large scale inflow rate towards the SMBH is given by the total
mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]),
\begin{eqnarray}
  \dot{M} &=& 4\pi \int_{0}^{r_{s}}q r^{2}dr = \frac{\eta
    M_{\star}|_{r_{\rm s}}}{\th} = \frac{\eta \Mbh (\rs/\rinf)^{2-\Gamma}}{\th} \nonumber \\
  &\approx&
  \begin{cases}
    4.5 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    3.2 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}            
\end{eqnarray}
where we have assumed $v_{w} \gg \sigma_0$ by adopting equation
(\ref{eq:stag_simple}) for $r_s$.  The resulting Eddington ratio is
given by
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    2.0 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core}, \\
    1.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
where $\dot{M}_{\rm edd} = 2.2M_{\bullet,8}M_{\odot}$ yr$^{-1}$ is the
Eddington accretion rate, assuming a radiative efficiency of ten per
cent.  Note the sensitive dependence of the inflow rate on the wind
heating rate.  Equation~\eqref{eq:eddr_analytic} is the radial mass
inflow rate on relatively large scales and does not include outflows
from the SMBH accretion disk (e.g.~\citealt{Blandford&Begelman99};
\citealt{Li+13}), which may significantly reduce the fraction of
$\dot{M}$ that actually reaches the SMBH. Thus we distinguish between
the large scale inflow rate, $\dot{M}$, and the accretion rate onto
the black hole, $\dot{M}_{\bullet}$.

@@

\begin{align}
  \frac{\dot{M}_{\rm TI}}{\dot{M}_{\rm edd}} \simeq \begin{cases}
    2\times 10^{-4}\eta_{0.02}^{0.47}M_{\bullet,8}^{0.36} &, \text{core}, \\
    8\times 10^{-5} \eta_{0.02}^{0.59} M_{\bullet,8}^{0.28}    &, \text{cusp},\\
  \end{cases}
  \label{eq:Mdotmax}
\end{align}

Note that since SMBH accretion rate cannot exceed the large scale
inflow rate, $\dot{M}_{\rm TI}$ also represents the maximum thermally
stable accretion rate.

What we describe above as ``thermal instability" may in practice
simply indicate an abrupt transition from a steady inflow-outflow
solution to a global cooling flow, as opposed a true thermal
instability.  In the former case the stagnation radius diverges to
large radii, increasing the density in the inner parts of the flow,
which increases cooling and creates a large inflow of cold gas towards
the nucleus.  A true thermal instability would likely result in a
portion of the hot ISM condensing into cold clouds, a situation which
may or may not be present in a cooling flow.  In this paper we do not
distinguish between these possibilities, although both are likely
present at some level.  Finally, note that if the CNM were to
``regulate" itself to a state of local marginal thermal instability
(as has previously been invoked on cluster scales; e.g.,
\citealt{Voit+15}), then equation (\ref{eq:Mdotmax}) might naturally
reflect the characteristic mass fall-out rate and concomitant star
formation rate.

@@

Key properties of the flow, such as the mass inflow rate and the
likelihood of thermal instability depend sensitively on the assumed
heating rate $\propto qv_{w}^{2}$.  In this section we estimate the
heating rate, $v_w$, taking into account contribution from stellar
winds, supernovae, millisecond pulsars and SMBH feedback. We assume
that all sources of energy injection are efficiently mixed into the
bulk of CNM gas. In fact, slower stellar wind material may cool and
form high density structures as in \citet{Cuadra+2005}. Additionally,
some of the injected energy may leak away from the bulk of the CNM
material through low density holes, as described by
\citet{Harper-Clark+2009} for stellar feedback and by
\citet{Zubovas+2014} for AGN feedback.

We summarize the individual heating sources below, providing estimates
of the value of the effective wind velocity
\begin{align}
  v_{w} = \sqrt{\frac{2 t_h \dot{e}}{\eta \rhostar}}
  \label{eq:vw_eff}
\end{align}
for each, where $\dot{e}$ is volumetric heating rate.

@@

where we implicitly the different sources of energy injection mix
efficiently. In reality, this may not be the case as slower velocity
sources cool and form high density structure due to high pressure from
the environment\citep{Cuadra+2005}.

@@

Perhaps a more readily addressable test is whether a steady-state
accretion picture developed in this work can account for the the large
scatter, typically of $2-3$ orders of magnitude, in $L_{X}/L_{\rm
  edd}$ at fixed $M_{\bullet}$.  Scatter could result from the strong
sensitivity of the accretion rate to the stellar wind velocity and
mass loss parameter, $\dot{M}/\dot{M}_{\rm edd} \propto
v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively
(Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).  When
combined with the significant dependence of $v_w$ on stochastic
intermittency in the star formation history (Appendix
\ref{fig:NickPlot2}), this can lead to order of magnitude differences
in $\dot{M}$. However, we note that $v_w$ is not expected to vary by
more than a factor of a few for a thermally-stable flow, limiting the
allowed variation of $\dot{M}/\dot{M}_{\rm edd}$. The discrete nature
of stellar wind sources and their motions could result in an
additional order of magnitude variation $\dot{M}$ \citep{Cuadra+2008}.

Variations in $L_{X}/L_{\rm edd}$ could also result from differences
in angular momentum of the infalling gas from galaxy to galaxy,
resulting in differences in the fraction of the gas lost to outflows.
Also potentially contributing is the order of magnitude difference, at
fixed $v_w$ and $M_{\bullet}$, between $\dot{M}$ for core and cusp
galaxies (Fig.~\ref{fig:bh_growth}).  Differences in $\dot{M}$ are
augmented by the theoretical expectation that $L_{\rm X} \propto
\dot{M}^{2} $ for radiatively inefficient flows.

@@

  \item The location of the stagnation radius, and hence the inflow
    rate, is a sensitive function of the gas heating rate $\propto
    v_w^{2}$.  Quantitatively, $\rs\propto v_w^{-2}$, implying that
    $\Mdot\propto v_w^{-3.8 (-2.4)}$ for fiducial core (cusp)
    galaxies.  However, $\Mdot$ can increase even more rapidly with
    decreasing $v_w$ for two reasons: (1) when $v_w$ becomes
    comparable to the stellar velocity dispersion ($\zeta < \zeta_c$;
    eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar
    bulge and cannot produce a flow with a positive kinetic energy
    interior to the stellar break radius, as we show in Appendix
    \ref{app:be}; (2) For low $v_w$, gas near the stagnation radius
    becomes thermally unstable, which based on the results of previous
    numerical studies (e.g.,\citealt{Ciotti+10}) is instead likely to
    result in a burst of high accretion ($\S\ref{sec:cycle}$).

@@

Finally, we do not account for the effects of non-spherical geometry
or discreteness of stellar wind sources \citep{Cuadra+2006,
  Cuadra+2008}. We assume all of the mass and energy sources are
efficiently mixed, and do not take into account the possibility of
slower wind material condensing into high density structures (as
described in \citealt{Cuadra+2005}) or some of the energy injection
leaking away in 3-d \citep{Harper-Clark+2009,Zubovas+2014}.
